Vulnerability Analysis of Container Images in Neuroimaging

### 2020-06-22 paper.tex
before structure modifications

Containers, which provide an operating system (OS) level virtualization, are
popular. This is because they are lightweight, flexible, resource-efficient,
provide fast development cycles, continuous delivery,
and cost reduction of infrastructure. OS-level virtualization gained a lot of interest when Docker was
introduced in 2013~\cite{gantikow2016providing}. However, Docker has security concerns due to which it
is not supported by High Performance
Computing (HPC) environments. As a result, Singularity came in 2016 with the goal of providing support
for HPC environments. It is used in many clusters including Compute Canada.
According to~\cite{kurtzer2017singularity}, Singularity is used in 36 research clusters.

@@

There are various image scanners available for scanning Docker images, for
instance Anchore, CoreOS's Clair, etc.
However, these scanners cannot be used to scan Singularity images due to different image formats.
Therefore, Singularity came up with Stools \href{https://github.com/singularityhub/stools}{(https://github.com/singularityhub/stools)}, which can be used for scanning Singularity images
for security and quality checks.

@@

This study involves several image scanning experiments. To begin with,
we scan the images as is to report the existing vulnerabilities
inside the images. Updating these images is likely to play a role in vulnerabilities. Therefore, we
update the images and then rescan all these
images.
Subsequently, we shrink the size of the images
meaning that remove unnecessary packages from the images to see the effect on the number of
vulnerabilities in the images.
Lastly, we update and shrink the image size together and then
rescan for the vulnerabilities.

In this section, we first
introduce two container tools used in our experiments: Docker and
Singularity and their image formats. Then we provide more insight
into the container images and the image scanners that we used in our study.
The image scanners are going to report different results due to different Linux
distributions/releases, which have different vulnerability data. Therefore, in the last, this section presents a
description of these varying Linux distributions/releases and their
vulnerability databases.

@@

\subsection{Container Tools}

In this paper, we consider two popular container tools i.e Docker and Singularity because
neuroimaging uses images from these two tools.
%Container tools are capable of packing and running an application in an isolated environment that has
%the application's dependencies,
%required libraries, and configuration files for the
%application to function properly.

\subsubsection{Docker}

@@

%consists of the following three main components: 
%
%\textit{Docker Daemon:} It acts as a server. It is
%a long-running process that manages all Docker images, containers, storage volumes, and networks.
%
%\textit{REST API:} It acts as an interface between the server and the client, and is an API which is used by the applications
%to talk to the \textit{Docker Daemon}.
%
%\textit{Client:} It is a Command Line Interface (CLI) client which is used to interact with the \textit{Docker Daemon}.
%Docker runs a software, \textit{Docker daemon}, on the host machine.
Docker Engine, which runs as root, is remotely controlled through a
UNIX socket. It is responsible for launching and managing containers,
consequently, when a container runs, the container is spawned as a
child process of this root-owned daemon.

\begin{comment}
It is responsible for launching containers, their
isolation, and spawning shells into the containers. It is also
responsible for managing Docker images i.e \textit{pull} and \textit{push}
from image repository, building images from Dockerfiles, signing them, etc.
When a Docker container runs, the container process is spawned as a
child of this root owned daemon. 
\end{comment}
As Docker users directly interact with the
Docker daemon through Command Line Interface (CLI) i.e \textit{docker} command, it is possible to force Docker daemon to escalate privileges.
If a user gets root privileges, it
introduces unthinkable security risks for multi-tenant shared
environments. Due to this reason, Docker is not supported by
HPC environments.

Docker images are organized as a series of layers stacked on top of each
other and identified by unique layer identifiers. If a layer is reused
in different images it will
have the same id across images. Each layer consists of a filesystem diff
with respect to the layer below.

\subsubsection{Singularity}

As Docker is not suitable for HPC systems, therefore, Singularity came up with a new approach
of packing and running applications in containers. 
In contrast to Docker,
Singularity does not run containers as root. Consequently, users cannot escalate
privileges in Singularity containers making it suitable for HPC environments.

A Singularity image is a single immutable file and
does not contain layers.
Once a Singularity image is created, it cannot be changed, consequently,
it can be used for reproducing the results of scientific
experiments.

@@

We scanned a total of 44 container images taken from the BIDS
apps~\cite{gorgolewski2017bids} and Boutiques~\cite{glatard2018boutiques}
frameworks, two containerization frameworks used in neuroscience to share
and re-use applications. Among the 27 BIDS apps \href{https://bids-apps.neuroimaging.io/apps/}
{(https://bids-apps.neuroimaging.io/apps/)} available at the time of
the study, we studied 26
unique Docker images as one wasn't available on DockerHub. Among the 49
Boutiques applications available at the time of the study, listed using
Boutiques \texttt{bosh search} command, 23 unique container images were
listed including 3 that couldn't be retrieved and 2 that were already
included in BIDS apps. The resulting 18 images included 12 Docker images
and 6 Singularity images.
%\begin{figure}
%\csvreader[%
% respect all,%
%  before reading=\footnotesize\sisetup{round-mode=places,round-precision=2,round-integer-to-decimal}
%      %\caption{Baseline GF Competition Circle Track 2014}\label{tab:baselline1}
%      \begin{adjustbox}{max width=\columnwidth},
%  after reading=\end{adjustbox},
% tabular={c |c | c | c | c},
%	table head =Image & {OS Distro} & {Vulnerabilities by Anchore} & {Vulnerabilities by Clair} & {Vulnerabilities by Vuls}\\\hline,
% late after line= \\,
%	late after last line=\\\hline %\multicolumn{5}{c}{Total Score: 1481}
%]{results.csv}{}{\csvlinetotablerow}%
%       \centering
%	\caption{Existing number of Vulnerabilities}
%\end{figure}

@@

Container image scanners can be broadly categorized into static and dynamic
scanners. Static scanning performs analysis on a image without running it
as a container. Therefore, package manager, which is a executable file,
cannot be used to extract required data while statically extracting image data.
Thus, scanners parse different files for extracting image data e.g, installed packages, their versions,
and OS distribution etc. For instance, in Debian and Ubuntu, \textit{/var/lib/dpkg/status} file
can be parsed to get a list of installed packages and in the same way \textit{/etc/os-release} file can be
parsed to get information about OS distribution.
Whereas in dynamic scanning, a container is deployed from the image first and then
package manager of the image is executed inside the container to collect required data.

@@

We used four popular image scanners, namely Anchore~\cite{github_2019},
Vuls~\cite{future-architect_2019}, Clair Scanner~\cite{arminc_2019} and
Stools~\cite{stools}. Anchore, Vuls and Clair Scanner support Docker
images while Stools supports Singularity. Vuls support both static as well as dynamic
scanning till version 0.9.5, however, in our experiments we used it dynamically.
On the contrary, the others
are static scanners. To avoid discrepancies, we froze the vulnerability
databases for these scanners on September 25, 2019.

@@

In our experiments, we used Vuls Docker image having version 0.9.0 with fast scan mode.
Vuls gets data from Open Vulnerability and Assessment Language (OVAL).
The interesting point is that there is no OVAL data for Ubuntu 17.04 and 17.10 distribution,
meaning that images with these distributions are impossible to scan with Vuls.

@@

Image scanners report vulnerabilities differently depending on the different Linux release types.
Therefore, we provide an overview of these Linux releases and scanners re. 
BIDS app and Boutiques images in our experiments have diverse Linux
distributions such as Ubuntu, Centos, Debian, and Alpine. Further, a particular
distribution have different releases in general. Here, we take Ubuntu's example
to discuss different releases and their vulnerability databases. We took Ubuntu's example
because mostly images in our experiments have this distibution. However, other distributions also have
similar release trend. In the case of Ubuntu, release
can be of two types: Regular or Long Term Support (LTS).
Regular release is supported only for 9 months, whereas LTS is supported for 5 years. Once LTS support
reaches End-Of-Life (EOL), Ubuntu decides whether to provide extended security support or not, which comes
in the form of Extended Security Maintenance (ESM) release.
ESM is a paid service that can be requested from Ubuntu to get security 
updates even after the EOL of a particular Ubuntu release. The ESM support period is decided by
Ubuntu. ESM and LTS releases of a particular Ubuntu distribution have different vulnerability
data in vulnerability databases, because after EOL of a release, a vulnerability is fixed only in packages of ESM, however, it still
exists in LTS.

@@

\textbf{released-esm:} The package (inside the given release) was vulnerable and
		an update has been already uploaded and published. However,
		this update is published in the Ubuntu ESM release only and not in LTS.
		The fixed version of such packages is appended by
		+esmN or \char`\~esm, indicating that this package version is available
		only via ESM.

@@

@@

On the other hand, if we want to reduce the number of vulnerabilities without effecting the reproducibility of the image,
image size reduction process is a good choice. The most obvious way of doing this is using 
Neurodocker (\href{https://github.com/ReproNim/neurodocker}{https://github.com/ReproNim/neurodocker}).
It traces and deletes unnecessary files.
However, image scanners reports vulnerabilities according to the installed
packages, whereas Neurodocker deals with the files only and list of installed packages remains same. Therefore, after
minimization also scanners report same number of vulnerabilities.
Hence, we need to reduce the number of packages
installed in the image by deleting unnecessary packages. For that to happen, we need to first figure
out which packages we need for the pipeline to function properly. We use
Reprozip~\cite{rampin2016reprozip}, a tool that
is used to make applications reproducible. It traces operating system calls used by the
application while execution and then based on that it recognizes binaries, files, environment variables,
and library dependencies that will be used by the application for future re-execution. If the OS is
Debian or RPM-based, Reprozip also
uses an available package manager to acquire a list of distribution packages from where these files come from.
So, we get a list of packages from Reprozip that are needed for the application execution. 
After that, we expand this list by adding important and required packages that are essential
for the system to function and finally figure out all the dependencies of these packages using Debtree
~\cite{debtree}.
Ultimately, we keep these packages and remove remaining unnecessary packages. We use bash scripts
(\href{https://github.com/kaurbhupinder/Vulnerability-Analysis/tree/master/Scripts/minification}{Github}) for
the image size reduction process, which installs Reprozip in the image, collects Reprozip trace, and
finally deletes unnecessary packages.

@@

\subsection{Existing Vulnerabilities}

@@

\subsection{Comparisons between Scanners}

Figure~\ref{fig:venn} represents the number of common and different vulnerabilities
that are shown by scanners.
When we add all regions of Venn diagram
we get total 20050 vulnerabilities in BIDS apps and Boutiques images.
It is interesting to note that out of these vulnerabilities, 4453 vulnerabilities are only
reported by Anchore, 673 reported only by Clair, and 325 detected only by Vuls.

>>>>

Containers, which provide an operating system (OS) level virtualization, are
popular. This is because they are lightweight, flexible, resource-efficient,
provide fast development cycles, continuous delivery,
and cost reduction of infrastructure. OS-level virtualization gained a lot of interest when Docker~\cite{docker19} was
introduced in 2013~\cite{gantikow2016providing}. However, Docker has security concerns due to which it
is not supported by High Performance
Computing (HPC) environments. As a result, Singularity~\cite{singularity19} came in 2016 with the goal of providing support
for HPC environments. It is used in many clusters including Compute Canada.
According to~\cite{kurtzer2017singularity}, Singularity is used in 36 research clusters.

@@

There are various image scanners available for scanning Docker images, for
instance Anchore, CoreOS's Clair, etc.
However, these scanners cannot be used to scan Singularity images due to different image formats.
Therefore, Singularity came up with Stools~\cite{stools}, which can be used for scanning Singularity images
for security and quality checks.

@@

This study involves several image scanning experiments. To begin with,
we scan the images as is to report the vulnerabilities
inside the images. Updating and shrinking the size of these images is likely to play a role in vulnerabilities. Therefore, we
update the images and then rescan all these
images.
Subsequently, we shrink the size of the images
meaning that remove unnecessary packages from the images to see the effect on the number of
vulnerabilities in the images.
Lastly, we update and shrink the image size together and then
rescan for the vulnerabilities.

In this section, we provide
insight
into the container images and the image scanners that we used in our study.
The image scanners refer to different Linux
distributions/releases, which have different vulnerability data and is likely to
create discrepancies in results. Therefore, in the last, this section presents a
description of these varying Linux distributions/releases and their
vulnerability databases.

@@

%\subsection{Container Tools}
%
%
%In this paper, we consider two popular container tools i.e Docker and Singularity because
%neuroimaging uses images from these two tools.
%Container tools are capable of packing and running an application in an isolated environment that has
%the application's dependencies,
%required libraries, and configuration files for the
%application to function properly.

%\subsubsection{Docker}

@@

%consists of the following three main components: 
%
%\textit{Docker Daemon:} It acts as a server. It is
%a long-running process that manages all Docker images, containers, storage volumes, and networks.
%
%\textit{REST API:} It acts as an interface between the server and the client, and is an API which is used by the applications
%to talk to the \textit{Docker Daemon}.
%
%\textit{Client:} It is a Command Line Interface (CLI) client which is used to interact with the \textit{Docker Daemon}.
%Docker runs a software, \textit{Docker daemon}, on the host machine.

%Docker Engine, which runs as root, is remotely controlled through a
%UNIX socket. It is responsible for launching and managing containers,
%consequently, when a container runs, the container is spawned as a
%child process of this root-owned daemon.
%
%\begin{comment}
%It is responsible for launching containers, their
%isolation, and spawning shells into the containers. It is also
%responsible for managing Docker images i.e \textit{pull} and \textit{push}
%from image repository, building images from Dockerfiles, signing them, etc.
%When a Docker container runs, the container process is spawned as a
%child of this root owned daemon. 
%\end{comment}
%As Docker users directly interact with the
%Docker daemon through Command Line Interface (CLI) i.e \textit{docker} command, it is possible to force Docker daemon to escalate privileges.
%If a user gets root privileges, it
%introduces unthinkable security risks for multi-tenant shared
%environments. Due to this reason, Docker is not supported by
%HPC environments.
%
%Docker images are organized as a series of layers stacked on top of each
%other and identified by unique layer identifiers. If a layer is reused
%in different images it will
%have the same id across images. Each layer consists of a filesystem diff
%with respect to the layer below.
%
%\subsubsection{Singularity}
%
%As Docker is not suitable for HPC systems, therefore, Singularity came up with a new approach
%of packing and running applications in containers. 
%In contrast to Docker,
%Singularity does not run containers as root. Consequently, users cannot escalate
%privileges in Singularity containers making it suitable for HPC environments.
%
%A Singularity image is a single immutable file and
%does not contain layers.
%Once a Singularity image is created, it cannot be changed, consequently,
%it can be used for reproducing the results of scientific
%experiments.

@@

We scan a total of 44 container images taken from the BIDS
apps~\cite{gorgolewski2017bids} and Boutiques~\cite{glatard2018boutiques}
frameworks, two containerization frameworks used in neuroscience to share
and re-use applications. Among the 27 BIDS apps \href{https://bids-apps.neuroimaging.io/apps/}
{(https://bids-apps.neuroimaging.io/apps/)} available at the time of
the study, we studied 26
unique Docker images as one wasn't available on DockerHub. Among the 49
Boutiques applications available at the time of the study, listed using
Boutiques \texttt{bosh search} command, 23 unique container images were
listed including 3 that couldn't be retrieved and 2 that were already
included in BIDS apps. The resulting 18 images included 12 Docker images
and 6 Singularity images.
%\begin{figure}
%\csvreader[%
% respect all,%
%  before reading=\footnotesize\sisetup{round-mode=places,round-precision=2,round-integer-to-decimal}
%      %\caption{Baseline GF Competition Circle Track 2014}\label{tab:baselline1}
%      \begin{adjustbox}{max width=\columnwidth},
%  after reading=\end{adjustbox},
% tabular={c |c | c | c | c},
%	table head =Image & {OS Distro} & {Vulnerabilities by Anchore} & {Vulnerabilities by Clair} & {Vulnerabilities by Vuls}\\\hline,
% late after line= \\,
%	late after last line=\\\hline %\multicolumn{5}{c}{Total Score: 1481}
%]{results.csv}{}{\csvlinetotablerow}%
%       \centering
%	\caption{Existing number of Vulnerabilities}
%\end{figure}

@@

%Container image scanners can be broadly categorized into static and dynamic
%scanners. Static scanning performs analysis on a image without running it
%as a container.

%Therefore, package manager, which is a executable file,
%cannot be used to extract required data while statically extracting image data.
%Thus, scanners parse different files for extracting image data e.g, installed packages, their versions,
%and OS distribution etc. For instance, in Debian and Ubuntu, \textit{/var/lib/dpkg/status} file
%can be parsed to get a list of installed packages and in the same way \textit{/etc/os-release} file can be
%parsed to get information about OS distribution.
%Whereas in dynamic scanning, a container is deployed from the image first and then
%package manager of the image is executed inside the container to collect required data.

@@

We used four popular image scanners, namely Anchore~\cite{github_2019},
Vuls~\cite{future-architect_2019}, Clair Scanner~\cite{arminc_2019} and
Stools~\cite{stools}. Anchore, Vuls and Clair Scanner support Docker
images while Stools supports Singularity. 
To avoid discrepancies, we froze the vulnerability
databases for these scanners on September 25, 2019.

%Vuls support both static as well as dynamic
%scanning till version 0.9.5, however, in our experiments we used it dynamically.
%On the contrary, the others
%are static scanners.

@@

In our experiments, we used Vuls Docker image having version 0.9.0 with fast scan mode.
Vuls gets data from Open Vulnerability and Assessment Language (OVAL).
The interesting point is that there is no OVAL data for Ubuntu 17.04 and 17.10 distribution,
meaning that images with these distributions are hard to scan with Vuls.

@@

Image scanners report vulnerabilities differently depending on the different Linux release types.
Therefore, we provide an overview of these Linux releases and scanners. 
BIDS app and Boutiques images in our experiments have diverse Linux
distributions such as Ubuntu, Centos, Debian, and Alpine. Further, a particular
distribution have different releases in general. Here, we take Ubuntu's example
to discuss its different releases and their vulnerability databases. We took Ubuntu's example
because mostly images in our experiments have this distibution. However, other distributions also have
similar release trend. In the case of Ubuntu, release
can be of two types: Regular or Long Term Support (LTS).
Regular release is supported only for 9 months, whereas LTS is supported for 5 years. Once LTS support
reaches End-Of-Life (EOL), Ubuntu decides whether to provide extended security support or not, which comes
in the form of Extended Security Maintenance (ESM) release.
ESM is a paid service that can be requested from Ubuntu to get security 
updates even after the EOL of a particular Ubuntu release. The ESM support period is decided by
Ubuntu. ESM and LTS releases of a particular Ubuntu distribution have different vulnerability
data in vulnerability databases, because after EOL of a release, a vulnerability is fixed only in packages of ESM, however, it still
exists in LTS.

@@

\textbf{released-esm:} The package (inside the given release) was vulnerable and
		an update has been already uploaded and published. However,
		this update is published in the Ubuntu ESM release only and not in LTS.
		The fixed version of such packages is appended by
		+esm or \char`\~esm, indicating that this package version is available
		only via ESM.

@@

%\begin{figure*} \centering{
%\includegraphics{status.pdf}}
%\caption{Experiment 1}
%\end{figure*}

%\includegraphics[width=300, max width=\textwidth]{status.pdf}

%\begin{table*}[!ht]
%\csvreader[%
% respect all,%
%  before reading=\footnotesize\sisetup{round-mode=places,round-precision=2,round-integer-to-decimal}
%      \begin{adjustbox}{max width=\textwidth},
%  after reading=\end{adjustbox},
% tabular={c |c | c | c | c | c | c | c| c | c},
%	table head =Image Labels & {Image names} & {Anchore} & {Clair} & {Vuls} &{Stools} & {a} & {b} & {c} & {d}\\\hline,
% late after line= \\,
%        late after last line=\\\hline %\multicolumn{5}{c}{Total Score: 1481}
%]{status.csv}{}{\csvlinetotablerow}%
%       \centering
%        \caption{\label{status}Number of vulnerabilities by scanners}
%\end{table*}

@@

On the other hand, if we want to reduce the number of vulnerabilities without effecting the reproducibility of the image,
image size reduction process is a good choice. The most obvious way of doing this is using 
Neurodocker~\cite{neurodocker}.
It traces and deletes unnecessary files.
However, image scanners reports vulnerabilities according to the installed
packages, whereas Neurodocker deals with the files only and list of installed packages remains same. Therefore, after
minimization also scanners report same number of vulnerabilities.
Hence, we need to reduce the number of packages
installed in the image by deleting unnecessary packages. For that to happen, we need to first figure
out which packages we need for the pipeline to function properly. We use
Reprozip~\cite{rampin2016reprozip}, a tool that
is used to make applications reproducible. It traces operating system calls used by the
application while execution and then based on that it recognizes binaries, files, environment variables,
and library dependencies that will be used by the application for future re-execution. If the OS is
Debian or RPM-based, Reprozip also
uses an available package manager to acquire a list of distribution packages from where these files come from.
So, we get a list of packages from Reprozip that are needed for the application execution. 
After that, we expand this list by adding important and required packages that are essential
for the system to function and finally figure out all the dependencies of these packages using Debtree
~\cite{debtree}.
Ultimately, we keep these packages and remove remaining unnecessary packages. We use bash scripts
(\href{https://github.com/kaurbhupinder/Vulnerability-Analysis/tree/master/Scripts/minification}{Github}) for
the image size reduction process, which installs Reprozip in the image, collects Reprozip trace, and
finally deletes unnecessary packages.

@@

\subsection{Detected Vulnerabilities}

@@

\subsection{Discrepancies between Scanners}

Figure~\ref{fig:venn} represents the number of common and different vulnerabilities
that are shown by scanners.

we get total 20050 vulnerabilities in BIDS apps and Boutiques images.
It is interesting to note that out of these vulnerabilities, 4453 vulnerabilities are only
reported by Anchore, 673 reported only by Clair, and 325 detected only by Vuls.
